---
title: "Crop Growth"
author: "Cooper Riggs, Adam Bitner, & Stockton Nelson"
format: pdf
editor: visual
---

## Setup

```{r}
#Downloading Libraries
library(tidyverse)
library(FRK)
library(spmodel)

#Bringing in Data
potato <- vroom::vroom("PotatoCWSI.csv") |>
  janitor::clean_names() |>
  na.omit()
```

## Libraries

-   tidyverse

-   FRK

-   spmodel

## Functions

```{r}
#Creating New Spatial Features
K <- 20
longlat <- as.matrix(potato[,1:2])
centers <- fields::cover.design(longlat, nd = K)$design
dist <- fields::rdist(centers)
scale <- 1.5 * max(apply(dist, 1, function(x){sort(x)[2]}))
spatial_features <- local_basis(
  manifold = plane(),
  loc = centers,
  scale = rep(scale, K),
  type = "bisquare") |>
  eval_basis(longlat) |>
  as.matrix()
colnames(spatial_features) <- paste0("SF", 1:K)
spotato <- bind_cols(potato, spatial_features)
```

```{r}
#Fitting the Spatial Model
psm <- splm(cwsi ~ slope + twi + aspect + eca_shallow + ndvi + SF1 + SF2 + SF3 + SF4 + SF5 + SF6 +SF7 + SF8 + SF9 + SF10 + SF11 + SF12 + SF13 + SF14 + SF15 +SF16 + SF17 + SF18 + SF19 + SF20, data = spotato, spcov_type = "exponential", xcoord = point_x, ycoord = point_y)

#Getting the Summary
summary(psm)
```

```{r}
#Finding the R^2
pseudoR2(psm)
```

## Cross Validation

```{r}
#Making a Cross Validation Function
set.seed(12)
cv <- function(fold){
  #Split the Data
  test  <- which(folds == fold)
  train <- setdiff(1:nrow(data), test)
  train <- data[train, ]
  test  <- data[test, ]
  
  #Fit the Linear Model
  model <- splm(cwsi ~ slope + twi + aspect + eca_shallow + ndvi + SF1 + SF2 + SF3 + SF4 + SF5 + SF6 +SF7 + SF8 + SF9 + SF10 + SF11 + SF12 + SF13 + SF14 + SF15 +SF16 + SF17 + SF18 + SF19 + SF20, data = train, spcov_type = "exponential", xcoord = point_x, ycoord = point_y)
  
  #Predict on the Test Set
  pred <- predict(model, newdata = test)
  
  #Compute RMSE
  sqrt(mean((test[["cwsi"]] - pred)^2))
}

#Setting Initial Values
data <- spotato
k <- 20
folds <- rep(1:k, length = nrow(data)) |> sample()

#Calculating the RMSE
sprmse <- sapply(1:k, FUN = cv)
mean(sprmse)
```

**RMSE:** \~0.3525

```{r}
#Making Fit and Prediction Data Frames 
preds_df <- vroom::vroom("PotatoCWSI.csv") |>
  janitor::clean_names() |>
  filter(is.na(cwsi))
preds <- preds_df |>
  select(point_x, point_y) |>
  as.matrix()
spatial_features_pred <- local_basis(
  manifold = plane(),
  loc = centers,
  scale = rep(scale, K),
  type = "bisquare") |>
  eval_basis(preds) |>
  as.matrix()
colnames(spatial_features_pred) <- paste0("SF", 1:K)
prediction_df <- bind_cols(preds_df, spatial_features_pred)

#Fitting the Spatial Model
ppsm <- splm(cwsi ~ slope + twi + aspect + eca_shallow + ndvi + SF1 + SF2 + SF3 + SF4 + SF5 + SF6 +SF7 + SF8 + SF9 + SF10 + SF11 + SF12 + SF13 + SF14 + SF15 +SF16 + SF17 + SF18 + SF19 + SF20, data = spotato, spcov_type = "exponential", xcoord = point_x, ycoord = point_y)

#Making Predictions Column
prediction_df$yhat <- predict(
  ppsm,
  newdata = prediction_df
)
```

```{r}
#Create a single object containing all the spatial feature info
spatial_info <- list(centers = centers,
                     scale = scale,
                     K = K)

#Saving Objects for Future Analysis
saveRDS(list(
  model = psm,
  spatial_info = spatial_info,
  data = preds_df), "cwsi_model.rds")
```

## Objects

-   Model: Spatial Model used for prediction on CWSI

-   Spatial_Info: Objects used to make the spatial features in the model

    -   centers: Points used as a reference for the spatial features.

    -   scale: Number which determines the reach of the spatial correlation.

    -   K: How many spatial features are created.

## Example Usage

```{r}

```
